\chapter{Introduction}
\label{chap:intro}

This blueprint documents the Lean~4 formalization of the combinatorial
aspects of associated cycles of special unipotent representations of
classical groups (Sp, SO, Mp). It follows two papers by
Barbasch--Ma--Sun--Zhu:

\begin{itemize}
\item \textbf{[BMSZb]} Barbasch--Ma--Sun--Zhu, \emph{Special unipotent
      representations: counting and reduction} (Section~10: PBP counting
      recursion formulas for classical groups).
      \href{https://arxiv.org/abs/2205.05266}{arXiv:2205.05266}.
\item \textbf{[BMSZ]} Barbasch--Ma--Sun--Zhu, \emph{Special unipotent
      representations: construction and unitarity} (Section~11:
      combinatorial associated cycles $\mathcal{L}_\tau$).
      \href{https://arxiv.org/abs/1712.05552}{arXiv:1712.05552}.
\end{itemize}

The formalization covers $\star \in \{B, D, C, M = \tilde{C}\}$; the real-form
variants $C^*$, $D^*$ are deliberately skipped (per a project directive).

\section{Group / dual-partition conventions}

For each root type $\star$ we work with a nilpotent orbit $\mathcal{O}$ in
the Lie algebra of a classical group. The \emph{BV dual} (Barbasch--Vogan /
Lusztig) $\check{\mathcal{O}}$ is a partition subject to parity conditions:

\begin{center}
\begin{tabular}{c|l|l}
  $\star$ & Group & Dual partition $\check{\mathcal{O}}$ \\\hline
  $B$ & $\mathrm{SO}(2n+1)$ & all parts even, positive \\
  $D$ & $\mathrm{SO}(2n)$   & all parts odd \\
  $C$ & $\mathrm{Sp}(2n)$   & all parts odd \\
  $M = \tilde{C}$ & $\mathrm{Mp}(2n)$ & all parts even, positive \\
\end{tabular}
\end{center}

From $\check{\mathcal{O}}$ we read off column lengths
$\mathbf{c}_1 \geq \mathbf{c}_2 \geq \cdots$ and row lengths
$\mathbf{r}_1 \geq \mathbf{r}_2 \geq \cdots$.

\section{Scope and structure}

This blueprint is organised as follows.

\begin{itemize}
\item Chapter~\ref{chap:pbp}: the painted bipartition (PBP) combinatorial data.
\item Chapter~\ref{chap:shape_shift}: shape-shifting bijections (Lemma~10.3 for M and C).
\item Chapter~\ref{chap:descent}: single/double descent (Lemmas~10.4, 10.5, Prop~10.8).
\item Chapter~\ref{chap:ddescent}: Prop~10.9 / Cor~10.10 double-descent injectivity.
\item Chapter~\ref{chap:counting}: main counting formulas (Prop~10.11/10.12).
\item Chapter~\ref{chap:ils}: Integer List Sequences, ILS operations (Section~9.4).
\item Chapter~\ref{chap:ac}: AC recursion Eq.~(11.2); sign/fcSign invariants.
\item Chapter~\ref{chap:tail}: tail PBP, Lemma~11.3, Prop~11.4.
\item Chapter~\ref{chap:11_5}: Lemma~11.5 two-step AC formula; Lemma~11.6 first entry.
\item Chapter~\ref{chap:11_main}: injectivity (11.8--11.13), main bijection (11.14/15/17).
\item Chapter~\ref{chap:status}: formal status.
\end{itemize}

All Lean declarations named via \texttt{\textbackslash lean\{...\}} can be
clicked to reach the source code in the GitHub repository.
