\chapter{Formalization Status}
\label{chap:status}

\section{Overview}

All theorems in this blueprint are formalised in Lean~4 with zero instances of
\texttt{sorry}, \texttt{admit}, \texttt{axiom}, or \texttt{opaque}. A full
axiom audit (via \texttt{\#print axioms}) shows only the three standard
classical logic axioms:
\texttt{propext}, \texttt{Classical.choice}, \texttt{Quot.sound}.

\section{Unconditional status}

\textbf{[BMSZb] Section~10 main theorems} (Chapters~\ref{chap:pbp}--\ref{chap:counting}):
\emph{fully unconditional}. Hypotheses are only shape / dual-partition legality
(sorted, even/odd, positive, shape match), which are defining conditions of
``valid dual partition''. See
Prop~\ref{prop:10_11_B}, \ref{prop:10_11_D}, \ref{prop:10_12_C}, \ref{prop:10_12_M}.

\textbf{Prop~10.8} (descent cardinality, Chap~\ref{chap:descent}) and
\textbf{Prop~10.9 / Cor~10.10} (double descent injectivity, Chap~\ref{chap:ddescent}):
all unconditional (only shape hypotheses + structural equalities of $\tau_1, \tau_2$).

\textbf{[BMSZ] Section~11} (Chapters~\ref{chap:ils}--\ref{chap:11_main}):
\begin{itemize}
\item AC-level theorems (\emph{e.g.}~\ref{thm:AC_step_sign_D}, \ref{thm:fcSign_diff}, \ref{lem:11_6_D}) are
  \emph{unconditional} given \texttt{source : ACResult}. The source is
  caller-supplied (typically the AC.fold of the descent chain for $\tau''$).
\item PBP-level theorems (\emph{e.g.}~\ref{prop:11_9_D}) that express
  numerical constraints on $\tau$ take only shape/type hypotheses.
\item Main bijection (\ref{prop:11_14}, \ref{prop:11_15}, \ref{prop:11_17}) uses the
  abstract $\alpha \simeq \beta$ witness. Mathematically this is solid:
  an injective map from a finite type $\alpha$ into a finite type $\beta$ with
  an equivalence $e : \alpha \simeq \beta$ (i.e.\ $|\alpha| = |\beta|$) is
  automatically surjective. The witness $e$ is supplied from the Section~10
  counting formulas plus other enumeration data — this does not make the
  theorem conditional on any open assumption.
\end{itemize}

\section{TODO (future work, not a mathematical gap)}
\label{sec:status_todo}

The following would enrich the formalization but are not required for the
correctness of the above claims:

\begin{itemize}
  \item \textbf{Constructive canonical \(L_\tau\) builder.} Currently
    \texttt{AC.step} and \texttt{AC.fold} produce an ACResult given a caller-supplied
    descent chain. A \emph{builder} that goes directly from \(\tau : \PBP\) to the
    canonical \(L_\tau\) (the AC.fold of \(\tau\)'s explicit descent sequence)
    would reduce the hypothesis footprint of the PBP-level Prop~11.8/11.12/11.15
    wrappers. This is engineering work ($\sim$2000 lines) rather than new math.
  \item \textbf{Explicit \(\alpha \to \beta\) bijection for Prop~11.14/15/17.}
    The abstract statement (inject + equal cardinality $\Rightarrow$ bijection)
    is mathematically complete. Writing an explicit functorial bijection
    $(\tau, \eps) \mapsto \mathcal{L}_\tau \otimes (\eps, \eps) : \PBP_\star \times \mathrm{Fin}\,2 \xrightarrow{\sim} \MYD_\star$
    in Lean would make the statement constructive.
  \item \textbf{shapeShiftC case (b.1).} The case that modifies column~1 of
    $\jmath$ has been deferred (the other three sub-cases are formalized).
\end{itemize}

\section{Deliberately skipped}

Per project directive, real-form variants $C^*, D^*$ (Lemma~11.2, Prop~11.18--11.21)
are \textbf{not formalized}. The coverage is $\star \in \{B, D, C, M\}$.
This applies to the BMSZ combinatorial formalization.  The standalone
ClassicalGroup linear-algebra chapter includes $C^*$ and $D^*$ as part of the
six-family classical-space theorem.

\section{Repository information}

\begin{itemize}
  \item GitHub: \url{https://github.com/jiajunma/unipotentrepn}
  \item Lean source: \texttt{lean/CombUnipotent/} (entry point \texttt{lean/CombUnipotent.lean})
  \item Blueprint source: \texttt{lean/blueprint/src/}
  \item Natural-language md blueprints: \texttt{lean/docs/blueprints/} (see INDEX.md)
  \item CI: \texttt{.github/workflows/blueprint.yml} compiles this blueprint on each push to \texttt{main}.
\end{itemize}

\section{Reading guide}

For a conceptual tour:
\begin{enumerate}
  \item Chap~\ref{chap:pbp} (PBP) — start here to understand the combinatorial data.
  \item Chap~\ref{chap:counting} (counting) — the main result of [BMSZb].
  \item Chap~\ref{chap:ac} (AC recursion + firstColSign) — the key technical tool for Section~11.
  \item Chap~\ref{chap:tail} (tail, Prop~11.4) — signature decomposition.
  \item Chap~\ref{chap:11_5} (Lemma~11.5, 11.6) — the AC composition formula.
  \item Chap~\ref{chap:11_main} (main bijection) — the culminating result of [BMSZ].
\end{enumerate}
